---
title: "Assessing the inequality in the political sentiments across the United Kinddom - A case study of #Indiref2"

author: |
  | `Author:`
  | `Adepeju, M.`
  | `Crime and Big Data Centre, Manchester Metropolitan University, Manchester, M15 6BH`
  
date: |
  | `Date:`
  | ``r Sys.Date()``

output:
  rmarkdown::html_vignette
  
#dev: png
#output:
  #word_document: default
  #always_allow_html: yes
#  pdf_document: default
always_allow_html: yes
#fig_caption: yes
##bibliography: references.bib

abstract: The xxx
  
vignette: >
  %\VignetteIndexEntry{Twitter-based analysis of political sentiments - A case of Brexit}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

<style type="text/css">

h1.title {
  font-size: 26px;
  line-height: 130%;
  color: Black;
  text-align: center;
}

h2.subtitle {
  font-size: 13px;
  line-height: 120%;
  color: Black;
  text-align: center;
}

h4.author { /* Header 4 - and the author and data headers use this too  */
  font-size: 17px;
  font-family: "Arial";
  color: Black;
  text-align: center;
}
h4.date { /* Header 4 - and the author and data headers use this too  */
  font-size: 17px;
  font-family: "Arial", Times, serif;
  color: Black;
  text-align: center;
}

h4.abstract { /* Header 4 - and the author and data headers use this too  */
  font-size: 10px;
  font-family: "Arial", Times, serif;
  color: black;
  text-align: center;
}

h4.institute{ /* Header 4 - and the author and data headers use this too  */
  font-size: 10px;
  font-family: "Arial", Times, serif;
  color: black;
  text-align: center;
}

body, td {
   font-size: 14px;
}
code.r{
  font-size: 13px;
}
pre {
  font-size: 13px
}
h1 { /* Header 1 */
  font-size: 16px;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 16px;
  color: DarkBlue;
}
h3 { /* Header 3 */
  font-size: 15px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;

</style>

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r functions, include=FALSE}
# A function for captioning and referencing images
fig <- local({
    i <- 0
    ref <- list()
    list(
        cap=function(refName, text) {
            i <<- i + 1
            ref[[refName]] <<- i
            paste("Figure ", i, ": ", text, sep="")
        },
        ref=function(refName) {
            ref[[refName]]
        })
})
```

# Introduction



The 2016 Brexit referendum has largely been framed as the United Kingdom’s voting to return sovereignty from Brussels to London. But in answering one nationalist call, the country has unleashed yet more nationalist forces that threaten to fracture its union.

 The United Kingdom is made up of four nations: England, Scotland, Wales, and Northern Ireland. In the three years since the vote to leave the European Union, Scotland’s independence movement, which was stymied by a failed referendum in 2014, has resurfaced. Calls for a similar poll in Northern Ireland, which would raise the prospect of reunification with the Republic of Ireland, have grown. Even in Wales—which, unlike Scotland and Northern Ireland, favored leaving the EU—nationalist sentiment appears to be increasing.







Brexiteers have largely branded Brexit as an independence movement—one which, after nearly half a century of EU membership, will enable Britain to reclaim control over its laws, its regulatory standards, and its trade. But the notion of transferring power from lawmakers in faraway capitals to local populations long predates Brexit here in Britain.

The promise of returned sovereignty was used by the British government, to some extent, to keep Scotland, Northern Ireland, and Wales in the union. All three have long held various powers of their own, but the process of “devolution” gathered speed in the late 1990s with the creation of legislative assemblies in Scotland (otherwise known as Holyrood), Northern Ireland (Stormont), and Wales (Senedd). Since the 2014 Scottish-independence referendum, further powers have been devolved, on region-specific issues including education, housing, and local tax.







More Stories



 Boris Johnson gestures as he speaks at the House of Commons.

Brexit and the Failure of Journalism 

Helen Lewis  



 People walk past pro-Brexit placards in central London.

How Britain Came to Accept a ‘No-Deal Brexit’ 

Helen Lewis  


 



 Anti-Brexit demonstrators attend a rally in central London.

This Is Just the Beginning of Brexit 

Tom McTague  











But now that the U.K. is leaving the EU—an outcome that was supported in England and Wales, but overwhelmingly rejected in Scotland and Northern Ireland—some have begun to question whether devolved powers alone are enough. Addressing reporters in Westminster this week, Scottish First Minister Nicola Sturgeon and Welsh First Minister Mark Drakeford, the leaders of their respective devolved governments in Edinburgh and Cardiff, expressed their frustrations with the British government’s newly renegotiated Brexit deal, and the lack of consideration the process has paid to their regional governments. While the deal was able to secure the preliminary support of a majority of lawmakers in the House of Commons on Wednesday, it has achieved little consensus elsewhere. Northern Ireland’s ruling Democratic Unionist Party declared the arrangement a “betrayal.” The devolved assembly in Wales rejected it outright. Scotland’s Parliament looks poised to do the same. (The deal does not, however, require the approval of these regional assemblies.)




“The uniqueness of this event—the first ministers of Wales and Scotland, of different political persuasions, uniting in opposition to this deal—is in itself a signal of how bad we believe it to be,” Sturgeon told reporters on Wednesday, noting that though Scotland and Wales came down on different sides of the Brexit vote, neither “voted for a hard Brexit … Nobody voted to have their core principles of devolution trampled over in a stampede to get out of the EU, come what may.”


Concern over the threat Brexit could have on the unity of the U.K. hasn’t been totally absent in the Brexit debate, though it’s largely been focused on Northern Ireland. The region’s legacy of sectarian fighting and the hard-won peace achieved by the 1998 Good Friday Agreement have complicated discussions of what a future land border between Northern Ireland, a part of the U.K., and the Republic of Ireland, a member of the EU, should look like. Both U.K. and EU negotiators sought to obviate the need for a customs border on the island of Ireland by reaching a compromise in which Northern Ireland would leave the EU customs union alongside Britain, but maintain closer alignment to EU rules than the rest of the United Kingdom. The DUP’s rejection of the deal was grounded in the fact that it regards any distinction between Belfast and the rest of the U.K. to be a threat to the union itself.




[Read: Brexit and Britain’s Northern Ireland déjà vu]

But whereas the DUP opposes this latest Brexit deal on the grounds that it gives Northern Ireland too much of a distinct status, Scottish leaders have rejected it because it doesn’t give Scotland the same thing. “While we support the Good Friday Agreement in its entirety and a deal for Northern Ireland that protects that, these arrangements would place Scotland at a serious competitive disadvantage,” Sturgeon said. She noted that while ample attention has been paid in Westminster to the interests of the country’s Brexit-supporting constituencies, “the fact that the nation of Scotland voted to remain is treated as irrelevant.” Indeed, a June poll by the British polling firm YouGov found that a majority of Conservative Party members said they would rather see Scotland and Northern Ireland break away from the rest of the U.K. if it meant ensuring that Brexit happens—a sentiment that has only furthered perceptions of Brexit as a project of English nationalism, rather than of Britain as a whole.













In 2014, a majority of the Scottish electorate voted to remain part of the United Kingdom. Since the Brexit vote, however, Scotland’s political calculus has changed: Support for Sturgeon’s Scottish Nationalist Party, which advocates for Scotland leaving the U.K., surged in European Parliament elections this year. Recent polls suggest that if another independence poll were held, the results would narrowly go in favor of secession.

Brexit isn’t the only factor fueling the independence movement. “Scottish attitudes towards Brexit have become mingled with Scottish opposition to the current government in Westminster,” Tom Devine, a Scottish historian and professor emeritus at the University of Edinburgh, told me.

That opposition has as much to do with the unpopularity of past Conservative governments in Scotland as it does with the party’s current leader, Prime Minister Boris Johnson. “He represents a lot of things that the Scots hate about modern Toryism,” Devine said, noting Johnson’s reputation for being privileged and out of touch. Johnson’s approval rating in Scotland is in negative double digits—a lack of popularity that, on his first visit to Edinburgh as prime minister, prompted him to leave his meeting with Sturgeon through the back door to avoid protesters.




But even if Brexit has raised the specter of the U.K. breaking apart, the sheer complexity of the past three years may yet save it. After all, if unraveling the decades-long relationship between Britain and the EU has proven daunting, what would disentangling centuries of Anglo-Scottish union look like?

[Read: Is Brexit worth Scotland’s independence?]

Devine said issues of security (an independent Scotland would pose challenges for Britain’s nuclear deterrent), borders (if Scotland were to rejoin the EU as the SNP has pledged, a hard border between Scotland and England would be required), and trade (the rest of the U.K. is Scotland’s largest trading partner) would almost certainly resurface. Referendum fatigue could also play a role.

“I wouldn’t necessarily bet on a successful independence vote given these factors,” he said, noting that while many Scots “in their hearts, in their emotions, [and] in their identity would want a separate Scotland, there are these problems … which of course will very much come to the fore during a campaign.”




Still, so long as Brexit is perceived less as a British project than as an English one, frustrations within the U.K.’s devolved regions are unlikely to go away. Drakeford, the Welsh first minister, warned that Westminster’s failure to consider the regions’ needs could risk disuniting the kingdom.

“We have different views of where the United Kingdom should end up,” Drakeford said. “But I’ve come to believe, over the months that I’ve had … meetings with the current government, that in the end it is the Unionists”—as the ruling Conservative Party is known—“who will see off the union.”

